MAESTROeX: A Massively Parallel Low Mach Number Astrophysical Solver

### 2019-06-06 MAESTROeX_I/results.tex
change line colors in figure 5 to match figure 4

To explore the accuracy and efficiency of the new temporal algorithm, we now analyze in detail three-dimensional, full-star calculations of convection preceding ignition in a sub-Chandrasekhar mass white dwarf. Again, we refer the reader to \cite{MAESTRO_AMR} and \cite{MAESTRO_convection} for setup details. We implement both uniformly- and irregularly-spaced base state with the new temporal algorithm, while only uniform base state spacing is used in the original algorithm. As in Section 3 of \cite{MAESTRO_AMR}, we choose the peak temperature and peak Mach number as the two diagnostics to compare the simulations. Figure \ref{fig:wdconvect_256_maxvar} shows the evolution of both peak temperature and peak Mach number until time of ignition on a single-level grid with resolution of $256^3$. The simulation using the new temporal scheme with uniformly-spaced base state gives the same qualitative results as the original scheme, and predicts very similar time of ignition ($t=7810$ s compared to $t=7850$ s for original algorithm). The simulation using the new temporal scheme with irregularly-spaced base state displays a slightly different peak temperature behavior during the initial transition period $t<150$ s, which results in the difference between the curves post transition. We strongly suspect that this is a result of using different initial model inputs for the star in which the resolution near the center of the star is much coarser with the irregular spacing than the uniform spacing. Fortunately, the simulation with irregular base state spacing still follows the same trend as with uniform spacing, and the star is shown to ignite at an earlier time $t=6840$ s.

Figure \ref{fig:wdconvect_Tmax} shows the peak temperature evolution over the first 1000 s on two grids of differing resolutions, $256^3$ and $512^3$. As previously suspected, the simulation with irregularly-spaced base state agrees much closer to the results from using uniform spacing as the resolution is increased. This is most likely due to the increased resolution of the initial star models for the irregular-spaced base state that is now able to properly capture the same properties near the center of the star as its uniformly-spaced counterpart. This is especially important when computing the base state pressure from base state density since we enforce a constant base state pressure outside of a prescribed cutoff density radius that could potentially be quite different with the two types of base state spacing.

@@

In the planar case, we use the same problem setup for a hot bubble rising in a white-dwarf environment as described in Section 6 of Paper V. Here we use a domain size of $3.6\times 10^7$ cm by $3.6\times 10^7$ cm by $2.88\times 10^8$ cm, and allow the grid structure to change with time. The single-level simulation at a resolution of $128^2 \times 1024$ was run on Cori Haswell with 48 processors and took approximately 33.5 s per time step (averaged over 10 time steps) using either the original or new temporal algorithm. The adaptive simulation has a resolution of $32^2 \times 256$ at the coarsest level, resulting in the same effective resolution at the finest level as the single-level simulation. We tag cells that satisfy $T-\bar{T} > 3\times 10^7$ K as well as all cell at that height. The adaptive run took only 3.7 s per time step, and this 89\% decrease in runtime is mostly due to the fact that initially only 6.25\% of the cells (1,048,576 out of $128^2 \times 1024$ cells) are refined at the finest level. Figure \ref{fig:bubble_results} shows a series of planar slices of the temperature profile at time intervals of 1.25 s, and verifies that the adaptive simulation is able to capture the same dynamics as the single-level simulation at much lower computational cost.

@@

We continue to use the full-star white dwarf problem described in Section \ref{sec:whitedwarf} to test adaptive simulations on spherical geometry. The adaptive grid is refined twice by tagging the density at $\rho > 5\times 10^7$ g cm$^{-3}$ on the first level and $\rho > 10^8$ g cm$^{-3}$ on the second level. The simulation on a single-level grid of $512^3$ resolution took 26.01 s per time step (again averaged over 10 time steps).  The adaptive grid has a resolution of $128^3$ at the coarsest level and an effective resolution of $512^3$ at the finest level. On this grid, 8.9\% of the cells (1,493,504 out of $256^3$ cells) are refined at the first level and 5.3\% (7,176,192 out of $512^3$ cells) at the second. The adaptive simulation took 11.55 s, resulting in more than a factor of 2 in speedup. Both simulations are computed to $t=400$ s and we choose to use the peak temperature as the diagnostics to compare the results. Figure \ref{fig:wdconvect_amr_Tmax} shows the evolution of the peak temperature for all three runs and shows that the adaptive simulation gives the same qualitative result as the single-level simulation. We do not expect the curves to match up exactly because the governing equations are highly nonlinear, and slight differences in the solution caused by solver tolerance and discretization error can change the details of the results. Each simulation was run on Cori KNL with 512 processors and 16 threads per core.

>>>>

To explore the accuracy and efficiency of the new temporal algorithm, we now analyze in detail three-dimensional, full-star calculations of convection preceding ignition in a sub-Chandrasekhar mass white dwarf. Again, we refer the reader to \cite{MAESTRO_AMR} and \cite{MAESTRO_convection} for setup details. We implement both uniformly- and irregularly-spaced base state with the new temporal algorithm, while only uniform base state spacing is used in the original algorithm. As in Section 3 of \cite{MAESTRO_AMR}, we choose the peak temperature and peak Mach number as the two diagnostics to compare the simulations. Figure \ref{fig:wdconvect_256_maxvar} shows the evolution of both peak temperature and peak Mach number until time of ignition on a single-level grid with resolution of $256^3$. The simulation using the new temporal scheme with uniformly-spaced base state gives the same qualitative results as the original scheme, and predicts a similar time of ignition ($t=7810$ s compared to $t=7850$ s for original algorithm). The simulation using the new temporal scheme with irregularly-spaced base state displays a slightly different peak temperature behavior during the initial transition period $t<150$ s, which results in the difference between the curves post transition. We strongly suspect that this is a result of using different initial model inputs for the star in which the resolution near the center of the star is much coarser with the irregular spacing than the uniform spacing. Fortunately, the simulation with irregular base state spacing still follows the same trend as with uniform spacing, and the star is shown to ignite at an earlier time $t=6840$ s.

Figure \ref{fig:wdconvect_Tmax} shows the peak temperature evolution over the first 1000 s on two grids of differing resolutions, $256^3$ and $512^3$. As previously suspected, the simulation using irregularly-spaced base state agrees much closer to the results from using uniform spacing as the resolution is increased. This is most likely due to the increased resolution of the initial star model for the irregular-spaced base state that is now able to properly capture the same properties near the center of the star as its uniformly-spaced counterpart. This is especially important when computing the base state pressure from base state density since we enforce a constant base state pressure outside of a prescribed cutoff density radius that could potentially be quite different with the two types of base state spacing.

@@

In the planar case, we use the same problem setup for a hot bubble rising in a white-dwarf environment as described in Section 6 of Paper V. Here we use a domain size of $3.6\times 10^7$ cm by $3.6\times 10^7$ cm by $2.88\times 10^8$ cm, and allow the grid structure to change with time. The single-level simulation at a resolution of $128^2 \times 1024$ was run on Cori Haswell with 48 processors and took approximately 33.5 s per time step (averaged over 10 time steps) using either the original or new temporal algorithm. The adaptive simulation has a resolution of $32^2 \times 256$ at the coarsest level, resulting in the same effective resolution at the finest level as the single-level simulation. We tag cells that satisfy $T-\bar{T} > 3\times 10^7$ K as well as all cells at that height. The adaptive run took only 3.7 s per time step, and this 89\% decrease in runtime is mostly due to the fact that initially only 6.25\% of the cells (1,048,576 out of $128^2 \times 1024$ cells) are refined at the finest level. Figure \ref{fig:bubble_results} shows a series of planar slices of the temperature profile at time intervals of 1.25 s, and verifies that the adaptive simulation is able to capture the same dynamics as the single-level simulation at much lower computational cost.

@@

We continue to use the full-star white dwarf problem described in Section \ref{sec:whitedwarf} to test adaptive simulations on spherical geometry. The adaptive grid is refined twice by tagging the density at $\rho > 5\times 10^7$ g cm$^{-3}$ on the first level and $\rho > 10^8$ g cm$^{-3}$ on the second level. The simulation on a single-level grid of $512^3$ resolution took 26.01 s per time step (again averaged over 10 time steps).  The adaptive grid has a resolution of $128^3$ at the coarsest level and an effective resolution of $512^3$ at the finest level. On this grid, 8.9\% of the cells (1,493,504 out of $256^3$ cells) are refined at the first level and 5.3\% (7,176,192 out of $512^3$ cells) at the second. The adaptive simulation took 11.55 s, resulting in more than a factor of 2 in speedup. Both simulations are computed to $t=400$ s and we choose to use the peak temperature as the diagnostic to compare the results. Figure \ref{fig:wdconvect_amr_Tmax} shows the evolution of the peak temperature for all three runs and shows that the adaptive simulation gives the same qualitative result as the single-level simulation. We do not expect the curves to match up exactly because the governing equations are highly nonlinear, and slight differences in the solution caused by solver tolerance and discretization error can change the details of the results. Each simulation was run on Cori KNL with 512 processors and 16 threads per core.
